%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% cubething.dev %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[leqno, 12pt]{article}
\documentclass[]{turabian-researchpaper}
\usepackage{../../cubething}

%opening
\title{Heidegger, Carnap, Husserl, and the Foundations of Rationality}
\author{Phoenix Ada Rose Mandala}

\begin{document}
	\maketitle

	\section{Heidegger against Husserl}

	Transcendentalist ego is itself too far removed from lived experience to be useful
	for a phenomenological investigation towards Being as such. We must proceed
	from deeper ends, namely our Being-in-the-world; our thrownness; our fundamental
	orientation towards the world in mood; etc. This is a \emph{hermeneutical}
	process because it involves an \emph{interpretation} of our lived experience.

	The goal of Husserl's investigations was epistemological - to find a firm basis
	for the sciences. But it is the history of philosophy - which gave rise to the
	sciences, and which praises them as the highest form of wisdom - which itself \emph{limits}
	the phenomenological method; its reliance on \emph{epistemology} and this
	field's reliance on Cartesian doubt - reconfigured in Husserl as the epoch\'e
	- its reliance on the ego as disconnected, as \emph{transcending} or \emph{ideal}
	- \emph{this} is what limits and dooms the Husserlian project. Only through an
	analysis of our \emph{factical, lived} experience will we be able to
	understand what drives us toward knowledge.

	Rorty says there is a gap left in the destruction of epistemology - this is
	where we \emph{discuss} - it is where the \textit{logos} happens. Carnap does not
	allow for fundamental ontology because of his dependence on a naturalistic,
	epistemological world-view. All of nature, reason, and logic is law-like,
	strict, non-contradictory. This is surely excellent for making predictions
	about the world, but it misses the way we actually live; it is a step \emph{beyond}
	the factical world and into transcendence. Metaphysics exists \emph{beyond} the
	epistemological. It exists prior to logic; it exists in \textit{logos}, in discussion,
	in interpretation, in dialogue or \emph{dialectics}. Knowledge proceeds in slow
	movement towards the true through contradiction and reconciliation - it is dialectical
	by nature. No epistemology with law-like characteristics - no naturalistic
	epistemology, no law of non-contradiction - can capture this.

	Symbolic formalism concretizes what is essentially a time-oriented process.
	Knowledge and limits do not exist outside of a time structure. The accumulation
	of knowledge is a slow \emph{progression}, and an \emph{unfolding} of truth. Truth
	may be eternal, but it is not \emph{static}. It depends on the time context
	and how truth is felt, seen, revealed to us is a matter of unfolding through
	time. It takes time even to formulate a thought.

	Science and math feel \textunderscore{static} in comparison to hermeneutics. Why?
	What is the distinction? This is contrasted to flow or wu-wei -- flow can
	happen in any case. There is a distinct static eternality to the universal.
	This is contrasted with the dynamic flow of the particular. The universal \emph{transcends}
	the particular and holds itself above in perfect stillness. What is the origin
	of this phenomenon?

	\section{On Kant}

	Transcendental ideas: God, the world, the soul - preconditions for rational knowledge;
	we cannot understand anything without presupposing these ideas. These cannot be
	\emph{known} because they exceed the boundaries of possible experience.

	Dialectical - to posit the limit is already to move beyond. The problem dissolves.
	Transcendental conditions of knowledge are revealed by dialectically opposed
	ideas; ideas which are rationally consistent but mutually exclusive. For
	example: freedom cannot be proven, but we act as if we are free; it is better if
	we act this way.

	"To make room for faith;" a belief in the absurd! - Connections to Kierkegaard.

	Todo: Draw on the Neokantian background for the origin of Husserlian
	phenomenology, the move away from Neokantianism towards the Vienna circle analytic
	thought (Marburg disillusionment with metaphysics, emphasis on physics,
	naturalism) and the Continental move towards Hegel (Southwest school, emphasis
	on history, humanities, dialectics).

	\section{Carnap against Heidegger}

	Carnap's conception of logic and of the world is constructionist. We build up
	the world from basic, phenomenological blocks, though these phenomenological
	building blocks are so minimal that they are hardly worth speaking of. For Carnap
	in the Aufbau, there is only a single building block. His criticism of Heidegger
	then can be clearly seen as an attack on Heidegger's violation of the logical basis
	of thought. It is an attack on psychologism, a trend which was deeply popular and,
	since Husserl, difficult to defend. Carnap is known to have been familiar with
	Husserl, even mentioning him in the Aufbau, and studying him closely before
	reformulating the Aufbau extensively around 19xx. However, Carnap, like many of
	his contemporaries, probably did not see the importance of the
	phenomenological project which underlies Husserl's prolegomena to the Logical
	Investigations. A quote from the excellent Introduction to Husserlain Phenomenology
	can elucidate this point.

	\begin{quote}
		If logic is naturally related to possible acts of thinking, then a precise, reciprocal
		demarcation of the conformity of such acts with logical as well as psychological
		laws is not merely the preparatory task of methodological work, rather it is
		the central task of logic itself. The rejection of logical "psychologism,"
		that is, the emancipation of "pure" logic from its ordination within (empricial-genetic)
		psychology, may not be separated from its positive counterpart, the cultivation
		of a novel, eidetic-descriptive "psychology." \autocite{bernet1993}
	\end{quote}

	Carnap, in rejecting psychologism, wishes to found our empirical understanding
	on as minimal a basis as possible. This is an admirable task. Heidegger, on the
	other hand, focuses heavily on the eidetic-descriptive task, outlining in detail
	various "modes of Being" under which we live our daily lives. For Heidegger, this
	is a primary task, it lies at a deeper root than logic can aspire to. For Carnap,
	a rich descriptive account of proto-logical thought is more or less what
	philosophers have already \emph{been} doing; it leads to confusion. Philosophy's
	task is to make life clear, and for Carnap this lies precisely in the logical-linguistic
	aspect of cognition. Heidegger's focus on emotive, poetic language is not only
	backwards, but impossible to formulate within a perfectly clear language. It
	is, strictly speaking, unthinkable, and therefore it is nonsense; worse, it is
	"bad poetry."

	\section{Metaphilosophical concerns}

	It would be helpful here to talk about The Task of Thinking. A summary:
	Philosophy as metaphysics - the study of the Being of beings - is nearing its
	end. It realizes this end in cybernetics, the total technoligization of human
	activity. By all accounts science is correct. What is left for philosophy? We have
	answered the call "to the things themselves," the call to examine subjectivity
	and its constitutive manifestations in full detail. we have received various answers
	and they have yielded from their speculation a totalizing science which itself
	subsumes the task philosophy set out to accomplish. But what remains is to
	understand where the underlying principles are grounded. What grounds rationality?
	What needs proof, and what does not? Heidegger speaks of a clearing, a space which
	opens for Lichtung to shine through. Lichtung is knowledge. But what is this clearing,
	what is it that shades and opens itself up so light can shine through? This is
	the task of thinking: the task of understanding \textit{al\=ethia} as unconcealment.

	Carnap, Husserl, Hegel and Heidegger himself were al misguided here. There is
	a "more originary questioning" which needs to occur.

	It's important to note the shared emphasis on science, logic, and technology
	betwene Hiedegger, Huserl, and Carnap. For Carnap, philosophy is the
	handmaiden of science. Philosophy's essential task was never metaphysics, but
	\emph{logic}. Epistemology is a part of this. Ontology plays no role; objects
	are constructed from simple sense perception and the rest follows a
	constructional systme. This is at least the view in the Aufbau; I do not
	believe he ever repudiated the main claim, that of constructionism. Heidegger
	says yes, you've got the idea: philosophy up to this point has been deeply
	focused on the Being of beings, in metaphysics. The end of metaphysics is
	science. To that end philosophy's goal - the end of philosophy - is cybernetcis,
	the full systemaization of human life. But there is a more oringary questioning.
	Logic is rationality emobdied - but logic itself admits of a psychological
	origin in phenomenology. The critique of psychologism cannot be complete without
	the admittance of a transcendental intuition, says Husserl; thus the sciences must
	be founded on phenomenology. Well and good says Heidegger, but let's dig deeper.
	What is it that allows us to be \emph{aware} of our intuition in the first place?
	On what grounds are the self-lighting, intuitive ideas made apparent? In a way,
	Heidegger is asking the hard problem: what is it that gives ground to
	consciousness? What grounds rationality?

	He does not give us an answer. To attempt that would be deeply immature. All we
	can hope for is a clarification of the question. This is what Heidegger hopes
	to give us, and it is this question itself which Heidegger believes all Western
	philosophy has until this point been blind to.

	--

	Heidegger's method relies on a historically grounded analysis of language, the
	development of language over time and an attempt to recreate the conditions of
	its origination in order to reveal the intention behind its usage. For Heidegger,
	this allows us to seek the underlying truths which are hidden behind layers of
	translation and historical obscurity. It allows us to see the foundational preconditions
	for philosophy as such, and so to form a critique of the foundational mode of
	philosophical knowledge. This is what is meant by the task of thinking: we are
	tasked with understanding how philosophy originated so that we can undergo
	this same process, perhaps to extend its purview or at least to expand the clearing
	by which we can see what really is. This is contrasted to Carnap's linguistic
	method, by which we attempt to get clear on concepts in a time-independent way.
	We abstract from language to an \emph{ideal} language, one which is perfect in
	its syntactical and semantic clarity. This is a wonderful approach, but one
	cannot make sense of it without the historical context in which it arose. A timeless,
	ideal language can only be conceptualized in the light of absolute Being, in light
	of the ideal. This conception is shaped by the Platonic Idea - this is
	apparent even in the name, "ideal." But what is an idea? From Descartes forward,
	an idea is a matter of subjectivity - so Carnap's goal is to find perfect
	clarity of rational mind: the objective principles by which subjectivity claims
	to understand the world. The basis of this questioning lies in the pursuit of
	\emph{truth.} But what is truth? Tarski would argue that truth is something
	system-dependent. Truth obtains within a logical system. The early Carnap and
	Heidegger, at which point the debate occurs, are both laboring under the
	notion of an absolute truth. It is unclear whether Carnap abandons this idea,
	but by the time he writes The Task of Thinking it is clear that Heidegger has
	moved beyond absolute truth. As explained above, Heidegger's pursuit has moved
	away from Being and Time toward Lighting and Clearing. What does he take truth
	to be?

	\begin{quote}
		Insofar as truth is understood in the traditionl "natural" sense as the correspondence
		of knowledge with beings, demonstrated in beings; but also insofar as truth is
		interpreted as the certainty of the knowledge of Being; \textit{al\=etheia},
		unconcealment in the sense of the clearing, may not be equated with truth.
		Rather, \textit{al\=etheia}, unconcealement thought as clearing, first
		grants the possibility of truth. For truth itself, like Being and thinking, can
		be what it is only in the element of the clearing. \autocite[446]{heidegger2008a}
	\end{quote}

	Here Heidegger is searching not only for the basis of the correspondence
	theory of truth, but also for the origin of fundamental ontology. He is
	looking for the basis of scientific thought as well as metaphysics. He takes truth
	to have multiple meanings, a meaning within science and a meaning within
	metaphysics. But how could it be that truth has multiple senses? There seems
	to be a rejection of the notion of absolute truth. Tarski would agree - and the
	later Carnap would famously propose a principle of tolerance, by which
	metaphysical confusion and confabulation can be replaced with a choice of a
	logical framework under which propositional statements (thus truth predicates)
	can be clearly and systematically articulated.

	\printbibliography
\end{document}