%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Professional Curriculum Vitae
% LaTeX Template
% Version 1.1 (9/12/12)
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
% Important note:
% This template requires the res.cls file to be in the same directory as the
% .tex file. The res.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{res}
% Use the res.cls style, the font size can be changed to 11pt or 12pt here

\usepackage{helvet} % Default font is the helvetica postscript font
%\usepackage{newcent} % To change the default font to the new century schoolbook postscript font uncomment this line and comment the one above
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{fontawesome5}
\usepackage{dirtytalk}

\newsectionwidth{0pt} % Stops section indenting

\begin{document}
	%----------------------------------------------------------------------------------------
	%	YOUR NAME AND ADDRESS(ES) SECTION
	%----------------------------------------------------------------------------------------

	\name{Phoenix Ada Rose Mandala\\ \hfill\textit{they/them}\hfill\\} % Your name at the top

	% If you don't want one of the addresses, simply remove all the text in the first or second \address{} bracket

	\address{ {\bf Contact} \\ \href{https://linkedin.com/in/ada-mandala}{\faIcon{linkedin} LinkedIn} \\
	%\href{https://cubething.dev/about}{\faIcon{globe} Personal Site} \\
	\href{https://github.com/ada-x64}{\faIcon{github} Github} \\ \href{mailto:phoenix.mandala@pm.me}{\faIcon{envelope} phoenix.mandala@pm.me} } % Your address 1
	\address{ {\bf Address} \\ 5800 Rock Chalk Drive, 15-104\\ Lawrence, KS 66049\\ \faIcon{phone} (785) 424-0182 \\ }

	%----------------------------------------------------------------------------------------

	\begin{resume}
		\setlength{\topsep}{0pt}
		\setlength{\parskip}{0pt}

		%----------------------------------------------------------------------------------------
		%	OBJECTIVE SECTION
		%----------------------------------------------------------------------------------------

		\section{\centerline{PROFESSIONAL SUMMARY}}

		\vspace{8pt} % Gap between title and text

		\textbf{Experienced Full-Stack Developer} with a strong background in transforming
		static designs into dynamic user experiences. My track record includes
		enhancing streaming services for Fortune 500 companies, developing game engines,
		and collaborating on a complex data visualization application. Passionate about
		\textbf{Rust, TypeScript, WASM, and UNIX}, with expertise in creating forward-thinking
		solutions across diverse technology stacks. Always looking to learn
		something new.

		\[
			\circ
		\]

		%----------------------------------------------------------------------------------------
		%	PROFESSIONAL EXPERIENCE SECTION
		%----------------------------------------------------------------------------------------

		\section{\centerline{PROFESSIONAL EXPERIENCE}}

		\vspace{8pt} % Gap between title and text

		{\sl Prospective} \hfill June 2023 -- June 2024 \\ Rust, WASM, Typescript, CI/CD,
		Python
		\begin{itemize}
			\itemsep -2pt

			\item Designed and developed the Prospective data visualization dashboard using
				the Perspective library, integrating with AWS, Azure, Google Cloud, and DigitalOcean.

			\item Created the plugin API and enhanced visual features; collaborated on
				JupyterLab extension, C++ data engine rewrite, and integrations with
				Kafka and Kerberos.
		\end{itemize}

		\vspace{4pt}
		{\sl Valorem Reply} \hfill June 2022 -- May 2023 \\ Rust, WASM, Python, Embedded
		Development
		\begin{itemize}
			\itemsep -2pt % Reduce space between items

			\item Contracted with Valorem Reply to bring ESPN+ to PlayStation 5 using
				a custom Rust framework.

			\item Developed infrastructure, implemented concurrent algorithms, and simplified
				analytics with a custom macro library; credited with \say{saving the project.}
		\end{itemize}

		\vspace{4pt}
		{\sl Roll20} \hfill October 2019 - August 2021 \\ Javascript, SCSS, HTML5, NodeJS
		\begin{itemize}
			\itemsep -2pt % Reduce space between items

			\item Developed front-end for user-driven IP BurnBryte, utilizing the
				Charactermancer feature and providing feedback to improve the API and
				developer experience.

			\item Created a local development environment mimicking the Roll20 frontend,
				boosting team productivity.

			\item Enhanced and extended the Starfinder character sheet code.
		\end{itemize}

		\vspace{4pt}
		{\sl Freelance} \hfill June 2018 - July 2022 \\ Javascript, PHP, C\#,
		Angular, React
		\begin{itemize}
			\itemsep -2pt % Reduce space between items

			\item Designed, developed, and integrated full-stack solutions to meet client
				expectations.

			\item Worked with clients on short-term contracts in various industries including
				game development, healthcare, and business applications.

			\item Built green and brownfield projects on web, mobile, and desktop.
		\end{itemize}

		\[
			\circ
		\]

		%----------------------------------------------------------------------------------------
		%	EDUCATION SECTION
		%----------------------------------------------------------------------------------------

		\section{\centerline{EDUCATION}}

		\vspace{8pt} % Gap between title and text

		{\sl BA}, Philosophy and Mathematics \\ University of Kansas \hfill December
		2014 \\ GPA 3.8

		\[
			\circ
		\]

		%----------------------------------------------------------------------------------------
		%	PROJECTS
		%----------------------------------------------------------------------------------------

		\section{\centerline{PERSONAL PROJECTS}}
		\vspace{8pt}
		{\sl Sundile - Game Engine} \hfill March 2022 - April 2023 \\ Rust, WASM, GPU
		\begin{itemize}
			\itemsep -2pt % Reduce space between items

			\item Developed a Rust-based graphics engine with WGPU for local and browser
				rendering. Utilizes WASM.

			\item Features include perspective cameras, .obj model rendering, and
				Blinn-Phong lighting.

			\item Implemented a custom asset engine for dynamic asset loading; project
				built from scratch for learning purposes.
		\end{itemize}

		\vspace{4pt}
		{\sl cubething.dev - Personal Site} \hfill February 2023 - Present\\ Deno,
		Preact, Server Side Rendering, CI/CD, DigitalOcean
		\begin{itemize}
			\itemsep -2pt % Reduce space between items

			\item Created a personal website utilizing Deno for the backend.

			\item Utilizes Preact to server-side render markdown blog documents.

			\item Continously deployed with Github Actions.
		\end{itemize}

		\[
			\circ
		\]

		%----------------------------------------------------------------------------------------
		%	SKILLS SECTION
		%----------------------------------------------------------------------------------------

		\section{\centerline{SKILLS}}

		\vspace{8pt} % Gap between title and text

		{\sl Programming and Templating Languages}
		\begin{itemize}
			\itemsep -2pt

			\item[] \textbf{Rust, Typescript, Javascript} (ECMAScript, Node, Bun, Deno),
				HTML5, CSS, SCSS, C, C++, C\#, PHP, LaTeX, CoQ, Lean, Markdown
		\end{itemize}

		\vspace{4pt}
		{\sl APIs and Stacks}
		\begin{itemize}
			\item[] \textbf{REST, MERN}, LAMP, Ruby on Rails, Flask, ASP.NET, Blazor
		\end{itemize}

		\vspace{4pt}
		{\sl Front-End / Client-Side }
		\begin{itemize}
			\item[] \textbf{WASM (Rust/C++), React, Browser APIs}, Preact, Yew, Lit,
				Backbone, Underscore, Web Components, LiveKit, Svelte
		\end{itemize}

		\vspace{4pt}
		{\sl Back-End / Server Side}
		\begin{itemize}
			\item[] \textbf{Fastify, Express.js, PostgreSQL, Nginx, Server-Side
				Rendering}, MySQL, Tower HTTP, SQLx, Axum
		\end{itemize}

		\vspace{4pt}
		{\sl Dev Tooling}
		\begin{itemize}
			\item[] \textbf{prettier, eslint, esbuild, git}, webpack, wasm-pack
		\end{itemize}

		\vspace{4pt}
		{\sl System Administration, Build Engineering, CI/CD}
		\begin{itemize}
			\item[] \textbf{Github Actions, Docker, NodeJS, Bun, zsh, bash, pwsh}, UNIX,
				Linux, MacOS, WSL2, Deno, pandoc, ripgrep, bat, sed, awk, Kafka
		\end{itemize}

		\vspace{4pt}
		{\sl Cloud Technologies}
		\begin{itemize}
			\item[] \textbf{Digital Ocean, PaaS, Serverless}, Microsoft Azure, AWS, Google
				Cloud
		\end{itemize}

		\vspace{4pt}
		{\sl GPU}
		\begin{itemize}
			\item[] \textbf{WebGPU, OpenGL, GLSL, HLSL}, Vulkan, three.js
		\end{itemize}

		\[
			\bullet
		\]
	\end{resume}
\end{document}